Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-22 paper/paper.Rmd
merged plots

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training. As a proof-of-concept for this idea, here we report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. This work provides an estimate of the probability that a motivated graduate student can reproduce a previously published finding of interest on a first attempt. Congruent with previous findings, replication studies typically yielded smaller effects than originals, though the modal outcome was some support for the original claim. We describe the workflow and pedagogical methods that were used in the class and discuss challenges for adoption of this pedagogical model and replication research more broadly."

@@

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
```

@@

It is a challenging proposition to estimate the replicability of an entire literature [cf. @osc2015]. Making such an estimate minimally requires determining a sampling strategy, a notion of what constitutes a direct replication, and a standard for what constitutes a successful replication, among many other difficult decisions [@gilbert2016;@anderson2016]. In the current work, we focus on a quantity that is easier to estimate: the probability that a graduate student can choose an article of interest in *Psychological Science* and -- in a single attempt, within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fidelity to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is reproducible in some more abstract sense.

@@

d$replicated_TA_factor <- factor(d$replicated_TA, levels = c(0, .5, 1), 
                                 labels = c("No","Partial","Yes"))
ggplot(d,aes(x=original_p_value_bins,y=rep_p_value, 
             color=d$replicated_TA_factor)) +
  geom_jitter(height=0, width=.2) +
  geom_hline(yintercept=c(0.05), lty = 2) + 
  geom_hline(yintercept=c(0.1), lty = 3) + 
  ylim(c(0,1)) +
  xlab("Original p-value (binned)") + 
  ylab("Replication p-value (exact)") + 
  theme_bw() + 
  scale_color_solarized(name = "Subjective replication rating") + 
  theme(legend.position = "bottom")
```

@@

```{r}
# Note that the sofer paper is off the charts (d > 16) and excluded
ggplot(d,aes(x=original_ES_d,y=rep_es, 
             color=d$replicated_TA_factor)) +
  geom_jitter(height=0, width=.2) +
  geom_vline(xintercept=c(0.2), lty = 5) +
  geom_vline(xintercept=c(0.5), lty = 2) +
  geom_vline(xintercept=c(0.8), lty = 3) +
  # geom_hline(yintercept=c(0.2), lty = 5) +
  # geom_hline(yintercept=c(0.5), lty = 2) +
  # geom_hline(yintercept=c(0.8), lty = 3) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  xlab("Original effect size (cohen's d)") + 
  ylab("Replication effect size (cohen's d)") + 
  theme_bw() + 
  scale_color_solarized(name = "Subjective replication rating") + 
  xlim(0, 4) +
  ylim(0, 4) +
  theme(legend.position = "bottom")

# Make beautiful ci chart here...

```

@@

```{r}
getBF <- function(t_stat, t_df) {
  return(ttest.tstat(t = t_stat, n1 = t_df, rscale = 1, simple = T))
};

@@

ggplot(tstatDF, aes(x = origBF, y = repBF)) +
  geom_point(alpha = .8, aes(size = df_ratio, 
                             col = replicated_TA_factor)) +
  geom_abline(slope = 1, lty = 2) + 
  geom_smooth(method = "lm", se=FALSE) +
  # geom_smooth(data = filter(tstatDF,
  #                           !(first_author_name %in% c("Atir","Ko"))),
  #                           method = "lm", se=FALSE, lty = 2, size = .5) +
  geom_text_repel(aes(label = first_author_name), size = 3) +
  xlab("Original Bayes Factor") + 
  ylab("Replication Bayes Factor") + 
  scale_x_log10(limits = c(10**-1.75, 10**11), 
                breaks = c(1/10, 1/3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1/10", "1/3", "3", "10","100", "1,000", 
                           "1,000,000","1,000,000,000")) +
  scale_y_log10(limits = c(10**-1.75, 10**11),
                breaks = c(1/10, 1/3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1/10", "1/3", "3", "10","100", "1,000", 
                           "1,000,000","1,000,000,000")) + 
  scale_size_continuous(name = "Ratio of replication df to original df") + 
  scale_color_solarized(name = "Subjective replication rating") + 
  theme_bw() +
  theme(legend.position="bottom")

# note: fix slope on trend line, because the atir finding is a hack right now so it fits in the axes. 
```

>>>>

abstract: 
    "Replications are important to science, but who will do them? One proposal is that students can conduct direct replications as part of their training. As a proof-of-concept for this idea, here we report a series of 11 pre-registered, direct replications of findings from the 2015 volume of Psychological Science, all conducted as part of a graduate-level course. This work provides an estimate of the probability that a motivated graduate student can reproduce a previously published finding of interest on a first attempt. Congruent with previous findings, replication studies typically yielded smaller effects than originals: The modal outcome was partial support for the original claim. We describe the workflow and pedagogical methods that were used in the class and discuss challenges for adoption of this pedagogical model and replication research more broadly."

@@

```{r global_options, include=FALSE}
rm(list=ls())
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=FALSE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(xtable)
library(readr)
library(tidyr)
library(ggplot2)
library(BayesFactor)
library(ggrepel)
#devtools::install_github("langcog/langcog")
library(langcog)
library(cowplot)
```

@@

It is a challenging proposition to estimate the replicability of an entire literature [cf. @osc2015]. Making such an estimate minimally requires determining a sampling strategy, a notion of what constitutes a direct replication, and a standard for what constitutes a successful replication, among many other difficult decisions [@gilbert2016;@anderson2016]. In the current work, we focus on a quantity that is easier to estimate: the probability that a graduate student can choose an article of interest in *Psychological Science* and -- in a single attempt, within constraints of budget, expertise, and effort -- reproduce the basic pattern of findings with sufficient fidelity to be able to build on it in future work. We believe that this notion of whether a piece of research supports cumulative progress is an important construct. Whether another independent scientist can build on a result is perhaps as important as -- and in many cases easier to measure -- whether the work is replicable in some more abstract sense.

@@

d$replicated_TA_factor <- factor(d$replicated_TA, levels = c(0, .5, 1), 
                                 labels = c("No","Partial","Yes"))
# ggplot(d,aes(x=original_p_value_bins,y=rep_p_value, 
#              color=d$replicated_TA_factor)) +
#   geom_jitter(height=0, width=.2) +
#   geom_hline(yintercept=c(0.05), lty = 2) + 
#   geom_hline(yintercept=c(0.1), lty = 3) + 
#   ylim(c(0,1)) +
#   xlab("Original p-value (binned)") + 
#   ylab("Replication p-value (exact)") + 
#   theme_bw() + 
#   scale_color_solarized(name = "Subjective replication rating") + 
#   theme(legend.position = "bottom")
```

@@

@@

```{r}
# Note that the sofer paper is off the charts (d > 16) and excluded
p1 <- ggplot(d,aes(x=original_ES_d,y=rep_es)) +
  geom_point(alpha = .8, 
             aes(color=replicated_TA_factor, size = collected_n)) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  scale_x_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  scale_y_continuous(limits = c(0, 3.5), 
                breaks = c(.2, .5, .8, 1, 2, 3)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_smooth(method = "lm", se=FALSE) +
  xlab("Orig. effect size (d)") + 
  ylab("Rep. effect size (d)") + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size(name = "Rep. N") +
  scale_color_solarized(name = "Subjective rep. rating")

getBF <- function(t_stat, t_df) {
  return(ttest.tstat(t = t_stat, n1 = t_df, rscale = 1, simple = T))
};

@@

p2 <- ggplot(tstatDF, aes(x = origBF, y = repBF)) +
  geom_point(alpha = .8, aes(size = df_ratio, 
                             col = replicated_TA_factor)) +
  geom_abline(slope = 1, lty = 2) + 
  geom_smooth(method = "lm", se=FALSE) +
  geom_text_repel(aes(label = first_author_name), size = 2, force = 2) +
  xlab("Orig. Bayes factor") + 
  ylab("Rep. Bayes factor") + 
  scale_x_log10(limits = c(10**-1.75, 10**11), 
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) +
  scale_y_log10(limits = c(10**-1.75, 10**11),
                breaks = c(1/10, .3, 3, 10, 100, 1e+3, 1e+6, 1e+9),
                labels = c("1e-1", "3e-1", "3", "1e1","1e2", "1e3", 
                           "1e6","1e9")) + 
  theme_bw(base_size=10) + theme(legend.position="bottom") + 
  scale_size_continuous(name = "Ratio of rep. df to orig. df") + 
  scale_color_solarized(name = "Subjective rep. rating")

# note: fix slope on trend line, because the atir finding is a hack right now so it fits in the axes.

plot_grid(p1, p2, labels = c("A", "B"))
```
